---
title: "Lecture 03 Example Code"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook illustrates functions for calculating descriptive statistics.

## Dependencies
This notebook requires the `ggplot2` package as well as the `janitor` and `skimr` packages. 

```{r load-packages}
# tidyverse packages
library(ggplot2) # for the mpg data

# other packages
library(skimr)   # descriptive statistics
```

We'll also use `base` and `stats`, but they are already loaded!

## Load Data
This notebook requires data from the `ggplot2` package:

```{r load-data}
auto <- mpg
```

## Illustrate Density Plots

First, we'll cover the new `geom` from `ggplot2` - `geom_density()`. This works the same way as other continuous geoms:

```{r density-geom}
ggplot(data = auto) + 
  geom_density(mapping = aes(x = hwy))
```

Remember to interpret this as density and not as the raw frequency of observations!

## Individual Descriptive Statistics

We can calculate mean, median, and standard deviation separately:

```{r individual-descriptives}
median(auto$hwy)
mean(auto$hwy)
sd(auto$hwy)
```

Note how this is specified (`df$var`) and how looks in terms of output - they aren't labeled. The output follows the sequence of the inputs - median, then mean, than mode. The average highway fuel efficiency is 23.440 miles per gallon and the typical deviation from this mean is plus or minus 5.955 miles per gallon - the bulk of observations lie between 17.4 and 29.4. Remember that, when there is missing data, all of these functions need the `na.rm = TRUE` optional argument!

The range can be found by subtracting the min value from the max:

```{r range}
max(auto$hwy)-min(auto$hwy)
```

The `range()` function will give you those max and min values rather than the difference:

```{r range-fxn}
range(auto$hwy)
```

## Summary Tables

We can get either very simple summaries or complex ones. The simple one is produced like so:

```{r simple-summary}
summary(auto$hwy)
```

We can also get a more complex summary:

```{r skim}
skim(auto)
```

Finally, remember that `skim()` can be used for just one or two variables as well:

```{r skim-selection}
skim(auto, hwy, cty)
```
